Wait-free Queues with Polygarithmic Step Complexity

### 2022-04-21 article/psuedocode.tex
apply april 20 feedbacks

\Statex $\diamondsuit$ \tt{\sl{Shared}}
\begin{itemize}
\item \tt{\sl{Node*} tree[]} \textsf{: A binary tree of \tt{Node}s such that \tt{Tree[0]} is the root and the left child and the right child and the parent of \tt{Tree[i]} are \tt{Tree[2i+1]}, \tt{Tree[2i+2]} and \tt{Tree[i/2]}.}
\end{itemize}

@@

\Statex $\blacktriangleright$ \tt{\sl{Node}}
\begin{itemize}
%\item \tt{\sl{*Node} left(), right(), parent()} \textsf{: computed as mentioned in the \tt{Tree[]}}
\item \tt{\sl{BlockList} blocks}\textsf{: Supports two operations \tt{blocks.tryAppend(Block b), blocks[i]}. Initially  empty, when \tt{blocks.tryAppend(b)} returns true \tt{b} is appended to the end of the list and \tt{blocks[i]} returns $i$th block in the \tt{blocks}. If some instance of \tt{blocks.tryAppend(b)} returns \tt{false} there is a concurrent instance of \tt{blocks.tryAppend(b$^\prime$)} which has returned \tt{true}.\tt{blocks[0]} contains an empty block with all fields equal to 0 and \tt{end} pointers to the first block of the corresponding children.}

@@

@@

\Statex
\Function{Object}{Dequeue()}{}
\State \tt{block newBlock= \Call{new}{\sl{LeafBlock}}} \cmt{Creates a block with null value element, appends it to the tree, computes its order among operations, then computes its response; if it exists returns the response's element.}
\State \tt{newBlock.element= null}
\State \tt{newBlock.sum\sub{enq}= tree[leaf].blocks[tree[leaf].head].sum\sub{enq}}
\State \tt{newBlock.sum\sub{deq}= tree[leaf].blocks[tree[leaf].head].sum\sub{deq}+1}
\State \tt{tree[leaf].}\Call{Append}{newBlock}
\State \tt{<r\sub{deq}, b\sub{r\sub{deq}}>=} \tt{tree[leaf].}\Call{Index}{tree[leaf].head, 1} \cmt{\tt{r} is the rank among all dequeues of the dequeue and \tt{b\sub{r\sub{deq}}} is the index of the block in the root containing the dequeue.}
\State \tt{<r\sub{enq}, b\sub{r\sub{enq}}>=} \Call{FindResponse}{r\sub{deq}, b\sub{r\sub{deq}}} \cmt{\tt{r\sub{deq}} is the rank of the enqueue which is the response to the dequeue or -1 if the response is \tt{null}.} \label{deqRest}
\If{\tt{r\sub{enq}==-1}}
\State \tt{output= null}
\State \tt{tree[0].blocks[b\sub{r\sub{deq}}].num\sub{finished}.inc()} \cmt{shared counter}
\If{\tt{tree[0].blocks[b\sub{r\sub{deq}}].num\sub{finished}==tree[0].blocks[b\sub{r\sub{deq}}].num}}\cmt{all the operations in the block containing the dequeue are finished.}
\State \tt{tree[leaf].last\sub{done}= b\sub{r\sub{deq}}}
\EndIf
\pagebreak
\Else
\State \tt{output= \Call{Get}{index\sub{response}}}
\State \tt{root.blocks[b\sub{r\sub{enq}}].num\sub{finished}.inc()}
\State \tt{root.blocks[b\sub{r\sub{enq}}].num\sub{finished}.inc()}
\If{\tt{root.blocks[b\sub{r}].num\sub{finished}==root.blocks[b\sub{r}].num}}
\State \tt{tree[leaf].last\sub{done}= b\sub{r\sub{enq}}}
\ElsIf {\tt{root.blocks[b\sub{r\sub{deq}}].num\sub{finished}==root.blocks[b\sub{r\sub{deq}}].num}} \cmt{\tt{root.blocks[b\sub{r}]} comes after \tt{root.blocks[b\sub{i}]}.}
\State \tt{tree[leaf].last\sub{done}= b\sub{r\sub{deq}}}
\EndIf
\EndIf
\State \Return{\tt{output}}
\EndFunction{Dequeue}

@@

\Function{int, int}{FindResponse}{int i, int b}\cmt{Computes the rank  and index of the block in the root of the enqueue that is the response of the \tt{i}th dequeue in the root's \tt{b}th block. Returns <-1,0> if the queue is empty.}
\If{\tt{root.blocks[b-1].size + root.blocks[b].num\sub{enq} - i $<$ 0}}
\State \Return \tt{-1, 0}
\Else
\State \tt{res= root.blocks[b-1].sum\sub{deq}- root.blocks[b-1].size + i}
\State \Return \tt{<res, tree[0].blocks.get(enq, res).order>}
\EndIf
\EndFunction{FindResponse}

@@

\Function{void}{Propagate()}{}
\If{\bf{not} \tt{this.}\Call{Refresh()}{}}
\State \tt{this.}\Call{Refresh()}{} \cmt{Lemma Double Refresh}
\EndIf
\If{\tt{this} \bf{is not} \tt{root}} \cmt{To check a node is the root we can check its index is 0.} 
\State \tt{this.parent().}\Call{Propagate()}{}
\EndIf
\EndFunction{Propagate}

@@

\Function{boolean}{Refresh()}{}
\State \tt{h= head}
%\If{\tt{n.blocks[h]!=null?}} \tt{h+=1}\EndIf
\State \tt{<new, n\sub{left}, n\sub{right}>= this.\Call{CreateBlock}{h}} \cmt{\tt{n\sub{left}, n\sub{right}} are the values read from the children's \tt{num\sub{propagated}}s.}
\If{\tt{new.num==0}} \Return{\tt{true}} \cmt{The block contains nothing.}
\ElsIf{\tt{root.blocks.tryAppend(new)}}
\ForEach{\tt{dir} {\keywordfont{in}} \tt{\{left, right\}}} \label{okcas}
\State \tt{\Call{CAS}{dir.super[n\sub{dir}], null, h+1}}
\State \tt{\Call{CAS}{dir.num\sub{propagated}, n\sub{dir}, n\sub{dir}+1}}
\EndFor
\State \tt{\Call{CAS}{head, h, h+1}}
\State \Return{\tt{true}}
\Else
\State \tt{\Call{CAS}{head, h, h+1}} \cmt{Even if another process wins, help to increase the head. It might fell sleep before increasing.}
\State \Return{ \tt{false}}
\EndIf
\EndFunction{Refresh}

@@

\Function{<Block, int, int>}{CreateBlock}{\sl{int} i} 
\Statex\cmt{Creates a block to be inserted into \tt{this.blocks[i]}. Returns the created block as well as values read from each childnum\sub{propagated} field. The values are used for incrementing children's num\sub{propagated}s if the block was appended to \tt{this.blocks} successfully.}
\State \tt{block newBlock= \Call{new}{\sl{block}}}
\State \tt{newBlock.group= num\sub{propagated}}
\State \tt{newBlock.order= i}
\ForEach{\tt{dir} {\keywordfont{in}} \tt{\{left, right\}}}
\State \tt{index\sub{last}= n.dir.head} \label{lastLine}
\State \tt{index\sub{prev}= n.blocks[i-1].end\sub{dir}} \label{prevLine}
\State \tt{block\sub{last}= n.dir.blocks[index\sub{last}]}
\State \tt{block\sub{prev}= n.dir.blocks[index\sub{prev}]}
\State \cmt{\tt{newBlock} includes \tt{n.dir.blocks[index\sub{prev}+1..index\sub{last}]}.}
\State \tt{n\sub{dir}= n.dir.num\sub{propagated}}
\State \tt{newBlock.end\sub{dir}= index\sub{last}}
\State \tt{newBlock.num\sub{enq-dir}= block\sub{last}.sum\sub{enq} - block\sub{prev}.sum\sub{enq}}
\State \tt{newBlock.num\sub{deq-dir}= block\sub{last}.sum\sub{deq} - block\sub{prev}.sum\sub{deq}}
\State \tt{newBlock.sum\sub{enq-dir}= n.blocks[i-1].sum\sub{enq-dir} + b.num\sub{enq-dir}}
\State \tt{newBlock.sum\sub{deq-dir}= n.blocks[i-1].sum\sub{deq-dir} + b.num\sub{deq-dir}}
\EndFor
%\State \tt{b.num\sub{enq}= b.num\sub{enq-left} + b.num\sub{enq-right}}
%\State \tt{b.num\sub{deq}= b.num\sub{deq-left} + b.num\sub{deq-right}}
%\State \tt{b.num= b.num\sub{enq} + b.num\sub{deq}}
%\State \tt{b.sum= n.blocks[i-1].sum + b.num}

\If{\tt{n} \bf{is} \tt{root}}
\State \tt{newBlock.size= max(root.blocks[i-1].size + b.num\sub{enq} - b.num\sub{deq}, 0)}
%\State \tt{b.sum\sub{non-null deq}= root.blocks[i-1].sum\sub{non-null deq} + max( b.num\sub{deq} - root.blocks[i-1].size - b.num\sub{enq}, 0)}
\EndIf

\State \Return \tt{b, c\sub{left}, c\sub{right}}
\EndFunction{CreateBlock}

@@

\Statex $\leadsto$ \textsf{Precondition:~\tt{n.blocks[b]} contains \tt{i}th enqueue in the node.}
\Function{element}{Get}{\sl{int} b, \sl{int} i} 
\If{\tt{this {\keywordfont is} leaf}} \Return \tt{this.blocks[b].element}
\Else
\If{\tt{i $\leq$ this.blocks[b].num\sub{enq-left}}} \cmt{\tt{i} exists in the left child of this node}
\State \tt{subBlock= this.left\Call{BSearch}{sum\sub{enq}, i, this.blocks[b-1].end\sub{left}+1, this.blocks[b].end\sub{left}}}
\State \Return\tt{this.left.}\Call{Get}{i-this.left.blocks[subBlock-1].sum\sub{enq}, subBlock} 
\Else
\State \tt{i= i-this.blocks[b].num\sub{enq-left}}
\State\tt{subBlock= this.right\Call{BSearch}{sum\sub{enq}, i, this.blocks[b-1].end\sub{right}+1, this.blocks[b].end\sub{right}}}
\State \Return\tt{this.right.}\Call{Get}{i-this.right.blocks[subBlock-1].sum\sub{enq}, subBlock} 
\EndIf
\EndIf
\EndFunction{Get}

\Statex
\Statex $\leadsto$ \textsf{Precondition: \tt{b}th block of the node has propagated up to the root and \tt{i}th dequeue in the node is in \tt{this.blocks[b]}.}
\Function{<int, int>}{Index}{\sl{int} b, \sl{int} i} \cmt{Returns the order in the root of $i$th dequeue in the $b$th block of node n among dequeues.}
\If{\tt{this {\keywordfont is} root}} \Return \tt{<i, b>}
\Else
\State \tt{dir= (this.parent().left==n)? left: right} \cmt{check \tt{n} is a left or a right child}
\State \tt{superBlock= \Call{BSearch}{this.parent(), this.sum\sub{deq-dir}, i, super[this.blocks[b].group]-p, super[this.blocks[b].group]+p}} \cmt{superblock's group has at most $p$ difference with the value stored in \tt{super[]}.}
\If{\tt{dir {\keywordfont is} left}}
\State \tt{i+= this.parent().blocks[superBlock-1].sum\sub{deq-right}}
\Else \State \tt{i+= this.parent.blocks[superBlock-1].sum\sub{deq} + this.blocks[superBlock].sum\sub{deq-left}} \cmt{consider dequeues from the right child}
\EndIf
\State \Return\Call{Index}{this.parent(), superBlock, i}
\EndIf
\EndFunction{Index}

@@

\Function{void}{Append}{\sl{block} b} 
\State \tt{head+=1} \label{appendEnd} \cmt{Lines \ref{appendStart} to \ref{appendEnd} are done by one process at time.}
\State \tt{b.group= head} \label{appendStart} \cmt{Append is only called by the owener of the leaf.}
\State \tt{blocks[head]= b} 
\State \tt{this.parent().}\Call{Propagate()}{} 
\EndFunction{Append}

@@

\Function{element}{Get}{\sl{int} i}
\cmt{Returns $i$th Enqueue.}
\State \tt{res= root.blocks.get(enq, i).order}
\State \Return{\Call{Get}{root, res, i-root.blocks[res-1].sum\sub{enq}}}
\EndFunction{Get}

>>>>

\Statex $\diamondsuit$ \tt{\sl{Shared}}
\begin{itemize}
\item \tt{\sl{Node*} tree[]} \textsf{: A binary tree of \tt{Node}s such that \tt{Tree[0]} is the root and the left child and the right child and the parent of \tt{Tree[i]} are \tt{Tree[2i+1]}, \tt{Tree[2i+2]} and \tt{Tree[i/2]}.} \label{treeRules}
\end{itemize}

@@

\Statex $\blacktriangleright$ \tt{\sl{Node}}
\begin{itemize}
\item \tt{\sl{*Node} left(), right(), parent()} \textsf{: initialized  when creating the tree as mentioned in the \tt{tree[]} description \ref{treeRules}.}
\item \tt{\sl{BlockList} blocks}\textsf{: Supports two operations \tt{blocks.tryAppend(Block b), blocks[i]}. Initially  empty, when \tt{blocks.tryAppend(b)} returns true \tt{b} is appended to the end of the list and \tt{blocks[i]} returns $i$th block in the \tt{blocks}. If some instance of \tt{blocks.tryAppend(b)} returns \tt{false} there is a concurrent instance of \tt{blocks.tryAppend(b$^\prime$)} which has returned \tt{true}.\tt{blocks[0]} contains an empty block with all fields equal to 0 and \tt{end} pointers to the first block of the corresponding children.}

@@

\begin{footnotesize}

\it{Conventions}
\begin{itemize}
  \item \tt{b\sub{x}}: the block containing \tt{x}
  \item \tt{r\sub{x}}: rank of \tt{x} in the current scope
\end{itemize}

\it{Unwritten rules}
\begin{itemize}
 \item \tt{blocks[b].sum\sub{x}=blocks[b].sum\sub{x-left}+blocks[b].sum\sub{x-right}}  \tt{ (for b>0 and x $\in$ \{enq, deq\}})
 \item \tt{blocks[b].sum=blocks[b].sum\sub{enq}+blocks[b].sum\sub{deq}}  \tt{ (for b>0})
  \item \tt{blocks[b].num\sub{x}=blocks[b].sum\sub{x}-blocks[b-1].sum\sub{x}} \\ \tt{(for b>0 and x $\in$ \{$\emptyset$, enq, deq, enq-left, enq-right, deq-left, deq-right\}})
\end{itemize}
\end{footnotesize}

@@

\Statex
\Function{Object}{Dequeue()}{}
\State \tt{block newBlock= \Call{new}{\sl{LeafBlock}}} \cmt{Creates a block with null value element, appends it to the tree, computes its order among operations, then computes its response; if it exists returns the response's element.}
\State \tt{newBlock.element= null}
\State \tt{newBlock.sum\sub{enq}= tree[leaf].blocks[tree[leaf].head].sum\sub{enq}}
\State \tt{newBlock.sum\sub{deq}= tree[leaf].blocks[tree[leaf].head].sum\sub{deq}+1}
\State \tt{tree[leaf].}\Call{Append}{newBlock}
\State \tt{<r\sub{deq}, b\sub{r\sub{deq}}>=} \tt{tree[leaf].}\Call{Index\sub{Deq}}{tree[leaf].head, 1}
\Statex \cmt{\tt{r} is the rank among all dequeues of the dequeue and \tt{b\sub{r\sub{deq}}} is the index of the block in the root containing the dequeue.}
\State \tt{<r\sub{enq}, b\sub{r\sub{enq}}>=} \Call{FindResponse}{r\sub{deq}, b\sub{r\sub{deq}}} 
\Statex \cmt{\tt{r\sub{enq}} is the rank of the enqueue which is the response to the dequeue or -1 if the response is \tt{null}.} \label{deqRest}
\If{\tt{r\sub{enq}==-1}}
\State \tt{output= null}
\State \tt{tree[0].blocks[b\sub{r\sub{deq}}].num\sub{finished}.inc()} \cmt{shared counter}
\If{\tt{tree[0].blocks[b\sub{r\sub{deq}}].num\sub{finished}==tree[0].blocks[b\sub{r\sub{deq}}].num}}\cmt{all the operations in the block containing the dequeue are finished.}
\State \tt{tree[leaf].last\sub{done}= b\sub{r\sub{deq}}}
\EndIf
\pagebreak
\Else
\State \tt{output= \Call{Get\sub{enq}}{r\sub{enq}, b\sub{r\sub{enq}}}}
\State \tt{root.blocks[b\sub{r\sub{enq}}].num\sub{finished}.inc()}
\State \tt{root.blocks[b\sub{r\sub{enq}}].num\sub{finished}.inc()}
\If{\tt{root.blocks[b\sub{r}].num\sub{finished}==root.blocks[b\sub{r}].num}}
\State \tt{tree[leaf].last\sub{done}= b\sub{r\sub{enq}}}
\ElsIf {\tt{root.blocks[b\sub{r\sub{deq}}].num\sub{finished}==root.blocks[b\sub{r\sub{deq}}].num}} \cmt{\tt{root.blocks[b\sub{r}]} comes after \tt{root.blocks[b\sub{i}]}.}
\State \tt{tree[leaf].last\sub{done}= b\sub{r\sub{deq}}}
\EndIf
\EndIf
\State \Return{\tt{output}}
\EndFunction{Dequeue}

@@

\Function{int, int}{FindResponse}{int i, int b}\cmt{Computes the rank  and index of the block in the root of the enqueue that is the response of the \tt{i}th dequeue in the root's \tt{b}th block. Returns <-1,0> if the queue is empty.}
\If{\tt{root.blocks[b-1].size + root.blocks[b].num\sub{enq} - i $<$ 0}}
\State \Return \tt{-1, 0}
\Else
\State \tt{response= root.blocks[b-1].sum\sub{deq}- root.blocks[b-1].size + i}
\State \Return \tt{<res, tree[0].blocks.get(enq, response).order>}
\EndIf
\EndFunction{FindResponse}

@@

\Function{void}{Propagate()}{}
\If{\bf{not} \tt{this.}\Call{Refresh()}{}}
\State \tt{this.}\Call{Refresh()}{} \cmt{Lemma Double Refresh}
\EndIf
\If{\tt{this} \bf{is not} \tt{root}} \cmt{To check a node is the root we can check its index is 0.} 
\State \tt{this.parent.}\Call{Propagate()}{}
\EndIf
\EndFunction{Propagate}

@@

\Function{boolean}{Refresh()}{}
\State \tt{h= head}
%\If{\tt{n.blocks[h]!=null?}} \tt{h+=1}\EndIf
\State \tt{<new, np\sub{left}, np\sub{right}>= this.\Call{CreateBlock}{h}} \cmt{\tt{np\sub{left}, np\sub{right}} are the values read from the children's \tt{num\sub{propagated}}s.}
\If{\tt{new.num==0}} \Return{\tt{true}} \cmt{The block contains nothing.}
\ElsIf{\tt{root.blocks.tryAppend(new)}}
\ForEach{\tt{dir} {\keywordfont{in}} \tt{\{left, right\}}} \label{okcas}
\State \tt{\Call{CAS}{dir.super[np\sub{dir}], null, h+1}}
\State \tt{\Call{CAS}{dir.num\sub{propagated}, np\sub{dir}, np\sub{dir}+1}}
\EndFor
\State \tt{\Call{CAS}{head, h, h+1}}
\State \Return{\tt{true}}
\Else
\State \tt{\Call{CAS}{head, h, h+1}} \cmt{Even if another process wins, help to increase the head. It might fell sleep before increasing.}
\State \Return{ \tt{false}}
\EndIf
\EndFunction{Refresh}

@@

\Function{<Block, int, int>}{CreateBlock}{\sl{int} i} 
\Statex\cmt{Creates a block to be inserted into \tt{this.blocks[i]}. Returns the created block as well as values read from each childnum\sub{propagated} field. The values are used for incrementing children's num\sub{propagated}s if the block was appended to \tt{this.blocks} successfully.}
\State \tt{block newBlock= \Call{new}{\sl{block}}}
\State \tt{newBlock.group= num\sub{propagated}}
\State \tt{newBlock.order= i}
\ForEach{\tt{dir} {\keywordfont{in}} \tt{\{left, right\}}}
\State \tt{index\sub{last}= dir.head} \label{lastLine}
\State \tt{index\sub{prev}= blocks[i-1].end\sub{dir}} \label{prevLine}
\State \tt{block\sub{last}= dir.blocks[index\sub{last}]}
\State \tt{block\sub{prev}= dir.blocks[index\sub{prev}]}
\State \cmt{\tt{newBlock} includes \tt{dir.blocks[index\sub{prev}+1..index\sub{last}]}.}
\State \tt{n\sub{dir}= dir.num\sub{propagated}}
\State \tt{newBlock.end\sub{dir}= index\sub{last}}
\State \tt{newBlock.sum\sub{enq-dir}= blocks[i-1].sum\sub{enq-dir} + block\sub{last}.sum\sub{enq} - block\sub{prev}.sum\sub{enq}}
\State \tt{newBlock.sum\sub{deq-dir}= blocks[i-1].sum\sub{deq-dir} + block\sub{last}.sum\sub{deq} - block\sub{prev}.sum\sub{deq}}
\EndFor
%\State \tt{b.num\sub{enq}= b.num\sub{enq-left} + b.num\sub{enq-right}}
%\State \tt{b.num\sub{deq}= b.num\sub{deq-left} + b.num\sub{deq-right}}
%\State \tt{b.num= b.num\sub{enq} + b.num\sub{deq}}
%\State \tt{b.sum= n.blocks[i-1].sum + b.num}

\If{\tt{this} \bf{is} \tt{root}}
\State \tt{newBlock.size= max(root.blocks[i-1].size + b.num\sub{enq} - b.num\sub{deq}, 0)}
%\State \tt{b.sum\sub{non-null deq}= root.blocks[i-1].sum\sub{non-null deq} + max( b.num\sub{deq} - root.blocks[i-1].size - b.num\sub{enq}, 0)}
\EndIf

\State \Return \tt{b, np\sub{left}, np\sub{right}}
\EndFunction{CreateBlock}

@@

\Statex $\leadsto$ \textsf{Precondition:~\tt{n.blocks[b]} contains \tt{i}th enqueue in the node.}
\Function{element}{Get\sub{Enq}}{\sl{int} b, \sl{int} i} 
\If{\tt{i $\leq$ blocks[b].num\sub{enq-left}}} \cmt{\tt{i} exists in the left child of this node}
\State \tt{subBlock= left\Call{BSearch}{sum\sub{enq}, i, blocks[b-1].end\sub{left}+1, blocks[b].end\sub{left}}}
\State \Return\tt{left.}\Call{Get}{i-left.blocks[subBlock-1].sum\sub{enq}, subBlock} 
\Else
\State \tt{i= i-blocks[b].num\sub{enq-left}}
\State\tt{subBlock= right\Call{BSearch}{sum\sub{enq}, i, blocks[b-1].end\sub{right}+1, blocks[b].end\sub{right}}}
\State \Return\tt{right.}\Call{Get}{i-right.blocks[subBlock-1].sum\sub{enq}, subBlock} 
\EndIf
\EndFunction{Get\sub{Enq}}

\Statex
\Statex $\leadsto$ \textsf{Precondition: \tt{b}th block of the node has propagated up to the root and \tt{i}th dequeue in the node is in \tt{blocks[b]}.}
\Function{<int, int>}{Index\sub{Deq}}{\sl{int} b, \sl{int} i} \cmt{Returns the order in the root of $i$th dequeue in the $b$th block of node n among dequeues.}
\State \tt{dir= (parent.left==n)? left: right} \cmt{check if a left or a right child}
\State \tt{superBlock= \Call{BSearch}{parent, sum\sub{deq-dir}, i, super[blocks[b].group]-p, super[blocks[b].group]+p}} 
\Statex\cmt{superblock's group has at most $p$ difference with the value stored in \tt{super[]}.}
\If{\tt{dir {\keywordfont is} left}}
\State \tt{i+= parent.blocks[superBlock-1].sum\sub{deq-right}}
\Else \State \tt{i+= parent.blocks[superBlock-1].sum\sub{deq} + blocks[superBlock].sum\sub{deq-left}} \cmt{consider dequeues from the right child}
\EndIf
\State \Return\Call{Index\sub{Deq}}{this.parent, superBlock, i}

\EndFunction{Index}

@@

\Function{void}{Append}{\sl{block} blk} 
\State \tt{head+=1} \label{appendEnd} \cmt{Lines \ref{appendStart} to \ref{appendEnd} are done by one process at time.}
\State \tt{blk.group= head} \label{appendStart} \cmt{Append is only called by the owener of the leaf.}
\State \tt{blocks[head]= blk} 
\State \tt{parent.}\Call{Propagate()}{} 
\EndFunction{Append}

\Function{element}{Get\sub{Enq}}{\sl{int} b, \sl{int} i} 
\State\Return \tt{blocks[b].element}
\EndFunction{Get\sub{Enq}}

@@

\Function{<int, int>}{Index\sub{Deq}}{\sl{int} b, \sl{int} i} 
\State\Return \tt{<i, b>}
\EndFunction{Index\sub{Deq}}
